### 2012-06-17 ras2012-paper/src/main/latex/paper/paper.tex
Re-evaluate solutions.

\documentclass[10pt,a4paper,draft]{article}
\usepackage[utf8x]{inputenc}
\usepackage{ucs}
\usepackage[english]{babel}
\usepackage{multirow}
\usepackage{rotating}
\author{Lukáš Petrovický}
\title{RAS2012: Competition Entry}
\begin{document}

\maketitle

\begin{abstract}
This article describes an entry to the 2012 RAS Problem Solving Competition, concerning dispatching on multi-track territories. The entry is based on the Drools Planner, a Java-based solver. On a reasonably recent computer, the resulting algorithm is able to produce feasible results within a minute and has been fine-tuned to provide best results in under 3 minutes. Source code to the entry is open source and well documented.
\end{abstract}

\section{Introduction}

\section{Describing the solution}

\section{Implementation}

\section{Achieved results}

Running the algorithm described above 20 times over each data set, we have compiled a set of results, see Table \ref{table:results}. All these were reached within 3 minutes in a single-threaded run, using Intel i7 Q820 processor running Fedora 17 and 2 GB of heap space inside Java 7 runtime environment.

\begin{table}
\caption{Solution performance per data set.}
\begin{tabular}{c||c|c|c|c|c|c}
\hline \hline
    & Best    & Q1      & Q2      & Q3      & Worst   & Std. dev. \\ 
\hline
TOY & \$808   & \$960   & \$1210  & \$1210  & \$1210  & 164 \\ 
\hline 
1   & \$1435  & \$1695  & \$1930  & \$2856  & \$5612  & 1219 \\ 
\hline 
2   & \$8097  & \$10364 & \$14865 & \$16177 & \$22486 & 4088 \\ 
\hline 
3   & \$10617 & \$13247 & \$15367 & \$17307 & \$21384 & 2986 \\ 
\end{tabular} 
\label{table:results}
\end{table}

Please note that via the benchmarking functionality of Drools Planner (see above), users may be able to fine-tune the algorithm to be focused either on providing better solutions or on faster turnaround times.

For statistics of the resolved systems, see tables \ref{table:TOY-808.tex}, \ref{table:RDS1-1435.tex}, \ref{table:RDS2-8097.tex} and \ref{table:RDS3-10617.tex} respectively.

\section{Conclusion}

\appendix

\section{Resolved systems}

In this section, we show the best solutions reached for each data set.

\include{TOY-808}
\include{RDS1-1435}
\include{RDS2-8097}
\include{RDS3-10617}

\section{Visualizations}

\begin{figure}
\centering
\includegraphics[width=150mm,angle=90]{solution.png}
\caption{RAS DATA SET TOY example territory. Undirected graph, arcs have descriptions.}
\end{figure}

\begin{figure}
\includegraphics[width=150mm,angle=90]{B1.png}
\centering
\caption{RAS DATA SET TOY example, route of Train B1. Directed graph where green marks the origin, red the destination and blue is where the train is allowed to wait.}
\end{figure}

In this section, we show examples of visualizations that the solution is capable of providing. These visualizations have been rendered on the fly using the JUNG library\footnote{Java Universal Network/Graph framework, http://jung.sourceforge.net}.

\end{document}